---
title: "SQL"
output: html_document
---

##Filtering Columns and Rows

With SQL, you can filter columns and rows by using SELECT and WHERE.  Let's look at an example using the Lahman database.  Wee should first load that database along with the package sqldf.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Type in the following to bring up categories
in front of you to view, which will help you 
know the data you are working with:
  
colnames(Batting)

Suppose now I would like to see the homerun totals for the 1927 Yankees, I could write the following query:

```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting 
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

To find all instances where Yankees have hit 
40 homeruns or more, use the following code:
  
```{r}
 query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='NYA' and HR>=30"
sqldf(query)
```

To find all instances where a player had 
more than 40HR but less than 60 strikeouts, 
first think thusly:  If I look at 
the options listed under "Batting" then I 
can assume strikeouts is SO.  Now type 
in the folliwng code:
  
```{r}
query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```

To find all instances of Phillies 
in the 1970s hitting more than 30 
homeruns, we know how to limit the 
team to Phillies and homeruns to greater 
than 30, but limiting the years can be 
difficult.  However, the years column 
contains numeric values like any other.  
We need to think of limiting years as 
limtiting any numeric value.  Hence type 
in the following code:
  
```{r}
 query<-"SELECT playerID,yearID,teamID,HR FROM Batting
WHERE teamID='PHI' and YEARID>=1970 and YEARID<1980 and HR>30"
sqldf(query)
```

## ORDER BY

With SQL, you can order the rows of your output.  Now suppose I want to find every instance of a player hitting more than 50 homeruns, and I want the players with the highest homerun total at the top.  I could write it thusly:

```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf (query)
```

If I want to find players who have fewer than 10 strikeouts in a season, yet more than 400 at-bats, I would type in the following:

```{r}
query<-"SELECT playerID, teamID, yearID, SO, AB  FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"

sqldf (query)
```


##Aggregation

Now let's take all of Bath Ruth's career homeruns and 
squash them into one total.  Use the following code:

```{r}
query<-"SELECT playerID, HR FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"

sqldf (query)
```


Now let's play with this a little further.
Use the following code to find the career homerun
totals for all players who have more than 600
total career homeruns:

```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf (query)
```

Now let's tackle this problem:  Find the players who have
averaged more than 30 homeruns per year throughout their
careers, with the highest averages at the top.  To do so,
type in the following:

```{r}
query<-"SELECT playerID, avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"

sqldf (query)
```

